---
title: "Will it Rain in Seattle?"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
editor_options:
  chunk_output_type: inline
---

Besides coffee, grunge and technology companies, one of the things that Seattle is most famous for is how often it rains. This dataset contains complete records of daily rainfall patterns from January 1st, 1948 to December 12, 2017.

* DATE = the date of the observation
* PRCP = the amount of precipitation, in inches
* TMAX = the maximum temperature for that day, in degrees Fahrenheit
* TMIN = the minimum temperature for that day, in degrees Fahrenheit
* RAIN = TRUE if rain was observed on that day, FALSE if it was not

Here are the packages that are being used
```{r}
library(ggplot2)
library(tidyverse) # 
library(dplyr)
library(reshape)
library(caret) # classification and regression training
library(pROC)

```

# Exploratory Data Analysis

Here we first read in and display the results of the file
```{r}
data.df <- read.csv("DataRainSeattle/seattleWeather_1948-2017.csv", header = TRUE)
head(data.df)
tail(data.df)

```

## Missing Data 
Taking care of missing data. Just removing the rows where the is no RAIN value recorded. Fixing the problem 

```{r}
which(is.na(data.df$RAIN))
data.df[which(is.na(data.df$RAIN)),]

# remove those rows
data.df <- data.df[-c(18416, 18417, 21068),]

which(is.na(data.df$TMAX))
which(is.na(data.df$TMIN))

data.df$DATE = as.Date(data.df$DATE)
data.df$RAIN <- as.factor(data.df$RAIN)
```
## Exploratory Analysis

```{r}
summary(data.df)

# plot of Rainy vs Non-Rainy days
ggplot(data.df, aes(RAIN)) + geom_bar() + labs(title="Count of Rainy Day",  subtitle="1948 - 2017")

```

```{r}

ggplot(data.df, aes(x=TMIN, y=PRCP)) + geom_point(size=2, shape=1) + xlab("Minimum Temperature")  + ylab("Precipitation (Inches)") + labs(title="Low Temperature vs Precip",  subtitle="1948 - 2017")



ggplot(data.df, aes(x=TMAX, y=PRCP)) + geom_point(size=2, shape=1) + xlab("Maximum Temperature")  + ylab("Precipitation (Inches)") + labs(title="Max Temperature vs Precip",  subtitle="1948 - 2017")


```



```{r}
data.df.melted = melt(data.df[,-5],id.vars = c("DATE"))

# grouped histogram plot

ggplot(data.df.melted, aes(x=variable, y=value, fill=value)) + geom_boxplot()


```







remove the outliners
```{r}

TMIN_todrop = which(data.df$TMIN < 17)
TMAX_todrop = which(data.df$TMAX > 97.5 | data.df$TMAX < 21.5)
PRECIP_todrop = which(data.df$PRECIP > 0.25 | data.df$PRECIP < -0.15)

index_todrop = union(TMIN_todrop, TMAX_todrop)
index_todrop = union(index_todrop, PRECIP_todrop)

data.df.filtered = data.df[-index_todrop,]

```

```{r}
#Function createDataPartition to create train and test dataset (0.8: 0.2)
index <- createDataPartition(data.df.filtered$RAIN, p = 0.8, list = FALSE)

# Training
train.df <- data.df.filtered[index,]

# Testing
test.df <- data.df.filtered[-index,]
```

Let's look at the summary of the results. We see that the mean values are aproximately the same, as well as the temperature TMAX and TMIN values. 

```{r}
head(train.df)
head(test.df)

summary(train.df)
summary(test.df)
```


# Training Model
```{r}
# glm logistic regression
# rain is being predicted as a function of the temperatures
model <- glm(RAIN ~ TMAX + TMIN, data = train.df, family = binomial)
summary(model)
```


```{r}
predicted_values <- predict(model, test.df[,-5], type = "response")
head(predicted_values)

```

# Validating Model
```{r}
# table of the test set
table(test.df$RAIN)
nrows_prediction<-nrow(test.df)

prediction <- data.frame(c(1:nrows_prediction))
colnames(prediction) <- c("RAIN")

prediction$RAIN <- as.character(prediction$RAIN)
prediction$RAIN <- "TRUE"


prediction$RAIN[ predicted_values < 0.5] <- "FALSE"
prediction$RAIN <- as.factor(prediction$RAIN)

table(prediction$RAIN, test.df$RAIN)
confusionMatrix(prediction$RAIN, test.df$RAIN)
```

```{r}

plot(roc(test.df$RAIN, predicted_values, direction="<"),
     col="cornflowerblue", lwd=3, main="ROC of the model", xlim=c(0.9,0),ylim=c(0,1.0))

# Call:
# roc.default(response = test.df$RAIN, predictor = predicted_values,     direction = "<")
# Data: predicted_values in 4394 controls (test.df$RAIN FALSE) < 3270 cases (test.df$RAIN TRUE).
#Area under the curve: 0.8282
```



